Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-25 paper/paper.tex
Add Craig interpolation reference
missing article

Program synthesis is the mechanised construction of software that provably
satisfies a given specification.  Synthesis tools promise to relieve the
programmer from thinking about \emph{how} the problem is to be solved;
instead, the programmer only provides a compact description of \emph{what}
is to be achieved.  Foundational research in this area has been
exceptionally fruitful, beginning with Alonzo Church's work on the
\emph{Circuit Synthesis Problem} in the sixties~\cite{church-synth}.
Algorithmic approaches to the problem have frequently been connected to
automated theorem proving~\cite{manna-waldinger,proof-planning}.

@@

Program synthesis is closely related to
superoptimisation, which is the problem of taking a
segment of code, usually written in C, and finding a minimal program which
has the same functionality.  The {\sc aha}~\cite{aha-paper} tool, GNU
superoptimiser~\cite{gnu-superoptimiser} and TOAST~\cite{toast} all perform superoptimisation of C
programs by enumerating possible programs and testing them on suitable
inputs.  This search can be compiled down to native code using an optimising
compiler, so for short programs the method is extremely fast.  However, this
method is impractical for synthesising longer programs because of the
exponential blowup of the search space.  We incorporate the core idea of
this technique by using explicit-state model checking in parallel with
symbolic model checking.

@@

 \begin{algorithmic}[1]
\Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{synth(inputs)}
    \If{candidate = unsatisfiable}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{verif(candidate)}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
\end{algorithm}

>>>>

Program synthesis is the mechanised construction of software that provably
satisfies a given specification.  Synthesis tools promise to relieve the
programmer from thinking about \emph{how} the problem is to be solved;
instead, the programmer only provides a compact description of \emph{what}
is to be achieved.  Foundational research in this area has been
exceptionally fruitful, beginning with Alonzo Church's work on the
\emph{Circuit Synthesis Problem} in the sixties~\cite{church-synth}.
Algorithmic approaches to the problem have frequently been connected to
automated theorem proving~\cite{manna-waldinger,proof-planning}.
Recent developments include an application of Craig interpolation
to synthesis~\cite{synth-interpolants}.

@@

Program synthesis is closely related to
superoptimisation, which is the problem of taking a
segment of code, usually written in C, and finding a minimal program which
has the same functionality.  The {\sc aha}~\cite{aha-paper} tool, the GNU
superoptimiser~\cite{gnu-superoptimiser} and TOAST~\cite{toast} all perform superoptimisation of C
programs by enumerating possible programs and testing them on suitable
inputs.  This search can be compiled down to native code using an optimising
compiler, so for short programs the method is extremely fast.  However, this
method is impractical for synthesising longer programs because of the
exponential blowup of the search space.  We incorporate the core idea of
this technique by using explicit-state model checking in parallel with
symbolic model checking.

@@

 \begin{algorithmic}[1]
\Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{\Call{synth}{inputs}}
    \If{candidate = unsatisfiable}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{\Call{verif}{candidate}}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
\end{algorithm}
